---
title: "MA214 Applied Statistics - Project 1 Week 2"
author: "Your Section and Group"
date: "`r Sys.Date()`"
output: 
  pdf_document:
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)

```

## Project 1 week2 worksheet

------------------------------------------------------------------------

### 1: Importing and Inspecting the Data

**Objective:** Load your dataset and understand its structure.

Things to consider:

-   What variables are available? Are there enough predictors for modeling?

-   What are their data types?

-   Are there missing values?

-   Do any variables need cleaning or recoding?

```{r, echo=TRUE}
library(ggplot2)
library(dplyr)
library(readr)
library(tidyr)
library(openintro)
data(evals)
# Import the data
# Example: data <- read.csv("your_dataset.csv")
# Example: data <- read.csv("URL_to_dataset.csv")

df<-evals


# View the first few rows
head(df)

# Check variable types and structure
str(df)

# Check for missing values
summary(df)

```

Group Discussion: Note any variables that need recoding or cleaning.

### 2. Data Transformation

**Objective:** Prepare your data for analysis by performing necessary transformations.

Consider:

-   Do variables need unit changes or standardization?

-   Would a log or other transformation improve visualization or interpretation?

-   Do you need to create new variables for analysis?

```{r, echo=TRUE}
# Perform any necessary transformations here
hist(df$cls_students)
hist(df$score)
hist(df$bty_avg)
hist(log((df$score)))
# Create new variables if needed
```

Group Discussion: Explain why each transformation was applied.

### 3. Data Summarization

**Objective:** Summarize your dataset to understand patterns and distributions.

Tasks:

-   Produce overall summaries using table, statistics

-   Produce grouped summaries if applicable (e.g., by category)

-   Produce meaningful summaries by plots

```{r, echo=TRUE}
# Overall summary
summary(df)

# Grouped summary example
df %>% group_by(ethnicity) %>% summarise(mean_value = mean(bty_avg, na.rm = TRUE))
df %>% group_by(gender) %>% summarise(mean_value = mean(bty_avg, na.rm = TRUE))

# Visual summary example
ggplot(df, aes(x=bty_avg)) + geom_histogram(binwidth=1)

numeric_vars <- df %>%
  select(score, bty_avg, age, cls_students, cls_perc_eval)

pairs(numeric_vars)

```

Group Discussion: Comment on interesting patterns observed in summaries and plots.

### 4. Assumption Checking

**Objective:** Evaluate key modeling assumptions using visualizations and summary statistics.

Check for:

-   Linearity

-   Normality

-   Independence

Tools: Histograms, boxplots, scatterplots

```{r, echo=TRUE}
# Check linearity and relationships
plot(df$bty_avg, df$score)+ geom_point()

# Histograms and boxplots
ggplot(df, aes(x=bty_avg)) + geom_histogram(binwidth=1)
ggplot(df, aes(y=bty_avg, x=gender)) + geom_boxplot()
ggplot(df, aes(y=bty_avg, x=ethnicity)) + geom_boxplot()

# Normality check
# qqnorm(data$Var)
# qqline(data$Var)
```

Group Discussion: Comment on whether assumptions are met and any potential issues.

### 5. Exploring Relationships Between Variables

Objective: Investigate associations and patterns between variables.

Questions:

Which variables appear associated?

Are relationships linear or nonlinear?

Are there outliers or unusual patterns?

```{r, echo=TRUE}
# Scatterplots or other visualizations
ggplot(df, aes(x=cls_students, y=bty_avg)) + geom_point()

# Correlation or other relationship measures
# cor(data$Var1, data$Var2, use="complete.obs")
```

Group Discussion: Summarize the key relationships and any patterns found.

------------------------------------------------------------------------

Submit this document at the end of today’s lab. If you downloaded any CSV files, please include them with your submission. **Each group should submit one set of files (Rmd, PDF, and any CSV files) to Blackboard.**

**To convert this R Markdown document to PDF:** In RStudio, click the **Knit** button at the top of the editor and select **Knit to PDF**. Make sure you have LaTeX installed (e.g., TinyTeX) to enable PDF generation.

| Name | BUID | Present/Absent |
|------|------|----------------|
|      |      |                |
|      |      |                |
|      |      |                |
|      |      |                |
|      |      |                |
